Spfy: an integrated graph-based sequence database for real-time prediction of Escherichia coli phenotypes

### 2018-02-08 paper-webserver.tex

No account creation is required to use the platform. A sharable token is automically created for users upon opening the website, and is embeded into the website address. Users can share results with their colleagues by copying the url, and files can be submitted for analysis from different computers with results reflected on each others' screens.

@@

TravisCI \url{travisci.io}, a continuous integration (CI) platform, is used to ensure that Spfy \url{https://github.com/superphy/backend} does not break with any changes to the codebase, and runs tests for functionality and backwards compatibility.
The individual tests use PyTest \url{https://doc.pytest.org/}, and are run within TravisCI's virtual environment. The current build status can be checked either on our Gtihub repository or at \url{https://travis-ci.org/superphy/backend}.
The CI is also used to automatically build Spfy's core Docker images, and upload them to Docker Hub \url{https://hub.docker.com/u/superphy/}.

@@

One of the key goals in developing Spfy was to accomadate a variety of result formats, from storing subtyping data to retrieving results as inputs for downstream analyses, such as statistical significance testing. We've shown how a graph database can accomdate a variety of programs used by bioinformaticians and is performant for data retrieval on the results of 10,000+ genomes. Spfy maintains instantaneity, as modern analytics platforms must respond to user-requested analyses of big-data in the same efficiency as old analyses of single files.

@@

% namely galaxy
Existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, and pipelines such as the Integrated Rapid Infectious Disease Analysis (IRIDA) platform \url{http://www.irida.ca/} and the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial} help automate the use of WGS data for public-health surveillance.
Reproducibility is important to science, and Galaxy aims to provide a reproducible, computation-based research environment which is accessible to individuals without programming knowledge. Galaxy defines a formal schema for linking different analysis software together, so analyses can be replicated end-to-end.
Building on top of the Galaxy framework, IRIDA adds prebuilt pipelines specific to bioinformatics uses, as well as sequence and result storage. IRIDA takes a project-based approach, with sequences stored per project, and results stored linearly per whole-genome sequence. Simiarily, BAP provides an integrated analysis pipeline for bacterial WGS data as a web service.
For result storage, these workflow technologies either use relational tables \cite{goecks2010galaxy}, or store resulting files to disk \cite{thomsen2016bacterial}.

Like IRIDA and BAP, Spfy automates workflows for users, and like Galaxy, Spfy uses task queues to distribute selected analysis.
% re: Matt "2. When comparing to IRIDA, BAP etc., can mention some differences with Superphy, e.g. the storage of interim result data that allows downstream integrated analysis"
On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
Spfy also processes XXXX files over XX tasks in XXXX time by using a novel approach of distributing computations over a task queue and multiple Docker compartmentalized containers.
We also focus on the data warehousing problem in bioinformatics: storing results as in IRIDA and BAP, but in a way where we can integrate results for further analyses.
Like BAP, Spfy intergrates different programs, such as for VF and AMR determination, into one pipeline. 
Because output from these programs is user-specific or transitory, results from identical comparisons are often recomputed. Additionally, output from different analyses are structured using distinct terminology and formats, which must be converted before they can be compared. Without a unified structure, these conversions quickly become impractical for broad usage. Graph-based storage of all results solves these problems.

@@

Future work will focus on adding additional analyses modules to aid genotype to phenotype predictions, and supporting different species. We're currently working on an approach using support vector machiens to predict \textit{E.coli} phenotypes \url{https://github.com/superphy/kmer} which will be integerated into Spfy. While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and can easily be integrated into existing platforms such as Galaxy \cite{goecks2010galaxy}, and IRIDA. Spfy's codebase is provided at \url{https://github.com/superphy/backend} and a developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.

>>>>

No account creation is required to use the platform. A sharable token is automatically created for users upon opening the website, and is embedded into the website address. Users can share results with their colleagues by copying the URL, and files can be submitted for analysis from different computers with results reflected on each others' screens.

@@

TravisCI \url{travisci.io}, a continuous integration (CI) platform, is used to ensure that Spfy \url{https://github.com/superphy/backend} does not break with any changes to the codebase, and runs tests for functionality and backwards compatibility.
The individual tests use PyTest \url{https://doc.pytest.org/}, and are run within TravisCI's virtual environment. The current build status can be checked either on our GitHub repository or at \url{https://travis-ci.org/superphy/backend}.
The CI is also used to automatically build Spfy's core Docker images, and upload them to Docker Hub \url{https://hub.docker.com/u/superphy/}.

@@

One of the key goals in developing Spfy was to accommodate a variety of result formats, from storing subtyping data to retrieving results as inputs for downstream analyses, such as statistical significance testing. We've shown how a graph database can accommodate a variety of programs used by bioinformaticians and is performant for data retrieval on the results of 10,000+ genomes. Spfy maintains instantaneity, as modern analytics platforms must respond to user-requested analyses of big-data in the same efficiency as old analyses of single files.

@@

% namely galaxy
Existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, and pipelines such as the Integrated Rapid Infectious Disease Analysis (IRIDA) platform \url{http://www.irida.ca/} and the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial} help automate the use of WGS data for public-health surveillance.
Reproducibility is important to science, and Galaxy aims to provide a reproducible, computation-based research environment which is accessible to individuals without programming knowledge. Galaxy defines a formal schema for linking different analysis software together, so analyses can be replicated end-to-end.
Building on top of the Galaxy framework, IRIDA adds prebuilt pipelines specific to bioinformatics uses, as well as sequence and result storage. IRIDA takes a project-based approach, with sequences stored per project, and results stored linearly per whole-genome sequence. Similarly, BAP provides an integrated analysis pipeline for bacterial WGS data as a web service.
For result storage, these workflow technologies either use relational tables \cite{goecks2010galaxy}, or store resulting files to disk \cite{thomsen2016bacterial}.

Like IRIDA and BAP, Spfy automates workflows for users, and like Galaxy, Spfy uses task queues to distribute selected analysis.
% re: Matt "2. When comparing to IRIDA, BAP etc., can mention some differences with Superphy, e.g. the storage of interim result data that allows downstream integrated analysis"
On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
Spfy also processes XXXX files over XX tasks in XXXX time by using a novel approach of distributing computations over a task queue and multiple Docker compartmentalized containers.
We also focus on the data warehousing problem in bioinformatics: storing results as in IRIDA and BAP, but in a way where we can integrate results for further analyses.
Like BAP, Spfy integrates different programs, such as for VF and AMR determination, into one pipeline. 
Because output from these programs is user-specific or transitory, results from identical comparisons are often recomputed. Additionally, output from different analyses are structured using distinct terminology and formats, which must be converted before they can be compared. Without a unified structure, these conversions quickly become impractical for broad usage. Graph-based storage of all results solves these problems.

@@

Future work will focus on adding additional analyses modules to aid genotype to phenotype predictions, and supporting different species. We're currently working on an approach using support vector machines to predict \textit{E.coli} phenotypes \url{https://github.com/superphy/kmer} which will be integrated into Spfy. While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and can easily be integrated into existing platforms such as Galaxy \cite{goecks2010galaxy}, and IRIDA. Spfy's codebase is provided at \url{https://github.com/superphy/backend} and a developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.
